---
title: "Predict Future Sales"
author: "Kazım Anıl Eren"
date: "`r Sys.Date()`"
output: 
  html_document:
    css: my_rmarkdown_theme.css
    fig_caption: yes
    number_sections: yes
    self_contained: no
    toc: no
---

```{r setup, include = FALSE, echo = FALSE, warning = FALSE, error = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  error = FALSE,
  results = 'asis',
  fig.align = 'center',
  fig.width = 10, 
  fig.height = 7
)
```

## Data Input ----
The aim is to predict sales for November 2015. Either train data will be aggregated to months / weeks  or the predictions will be based on daily data (and then aggregated for the month of course).

```{r Library Input}
rm(list = ls()); gc(); closeAllConnections();
source("C:/Users/kazimanil/Documents/Digitallency_GGPlot_Theme.R")
```

```{r Initial Data Input}
train = fread("data/sales_train_v2.csv")[, .(date = as.Date(date, format = "%d.%m.%Y"), shop_id, item_id, item_price, sale = item_cnt_day)]
test  = fread("data/test.csv")
nov   = data.table(date = seq(as.Date("2015-11-01"), as.Date("2015-11-30"), 1))
subm  = fread("data/sample_submission.csv")
shop  = fread("data/shops-translated.csv", encoding = "UTF-8")
item  = fread("data/item_category.csv", encoding = "UTF-8")
curr  = fread("data/usd-rub.csv")
cal   = fread("data/calendar.csv")
dates = merge(cal,
              curr, 
              by = "date")
rm(cal, curr)
```

Test set is on a monthly basis. So my estimations should be based on monthly sales as well. I will focus on the months which have more than 0 sales.

```{r Aggregation by Shop-Month}
train_monthly = train[, .(saleQ = sum(sale), saleR = sum(sale * item_price))
                      , .(year = year(date), month = month(date), shop_id, item_id)]
train_monthly = train_monthly[saleQ > 0]
train_monthly = train_monthly[order(item_id, shop_id, year, month)]
train_monthly[, saleP := saleR / saleQ] # Sale price is Revenue divided by Quantity.
```

```{r Histogram of Sales, fig.cap = "Sale Quantities of Products by Month and Shop"}
ggplot(data = train_monthly, aes(x = saleQ)) + 
  geom_histogram(binwidth = 1, col = "orange") +
  theme_dt() + labs(x = "Sale Quantity", y = "Frequency")
```

There is a huge amount of observations in minor values. Thus, let's have a closer look. I will limit the monthly sales by 20 now:

```{r Histogram of Sales - Filtered, fig.cap = "Sale Quantities of Products by Month and Shop - Filtered"}
ggplot(data = train_monthly[saleQ < 20], aes(x = saleQ)) + 
  geom_histogram(binwidth = 1, col = "orange") +
  theme_dt() + labs(x = "Sale Quantity", y = "Frequency")
```

Second histogram hints out a o possible *Poisson distribution* of sales among shops and products. So better be careful while using a linear regression! For now, let's forget about the estimation progress but rather focus on building on our **Exploratory Data Analysis**. I will now, aggregate the on Shop-Month level and merge the dataset with my shop specifications dataset derived from shop names. Then will have some plots as seen below:

```{r Aggregation by Shop}
train_shop = train_monthly[, .(Revenue = sum(saleR)), .(shop_id, year, month)]
train_shop = train_shop[order(shop_id, year, month)]
train_shop[, Date := as.Date(paste0(year, "-", month, "-01"))]
train_shop = merge(train_shop, shop, by = "shop_id")
train_shop[City == "Moscow", LocType := "Moscow"]
train_shop[City %in% c("Web", "OutboundTrade"), LocType := "Web"]
train_shop[is.na(LocType), LocType := "Other"]
```

```{r Shop Type Graph, fig.cap = "Monthly Average Revenue by Shop Types"}
data = train_shop[, .(RevSum = sum(Revenue), ShopAmount = .N, AvgRev = sum(Revenue) / .N), .(Type, year, month)]
data[, Date := as.Date(paste0(year, "-", month, "-01"))]
ggplot(data, aes(x = Date, y = AvgRev / 1000000, col = Type)) +
  theme_dt() + labs(x = "Date", y = "Average Revenue (in Million Rubles)", title = "Monthly Average Revenue by Shop Types") +
  # geom_smooth(se = FALSE, size = 2) + 
  geom_line(size = 1) + 
  scale_x_date(date_breaks = "3 months",  date_labels = "%b %y") +
  theme(legend.position = "right", legend.title = element_text("Shop Types"))
```

A smoothed-out graph (using LOESS) is more comprehensible:

```{r Shop Type Graph Smoothed-Out, fig.cap = "Monthly Average Revenue by Shop Types - Smoothed Out"}
data = train_shop[, .(RevSum = sum(Revenue), ShopAmount = .N, AvgRev = sum(Revenue) / .N), .(Type, year, month)]
data[, Date := as.Date(paste0(year, "-", month, "-01"))]
ggplot(data, aes(x = Date, y = AvgRev / 1000000, col = Type)) +
  theme_dt() + labs(x = "Date", y = "Average Revenue (in Million Rubles)", title = "Monthly Average Revenue by Shop Types - Smoothed Out") +
  geom_smooth(se = FALSE, size = 2) + #geom_line(size = 1) + 
  scale_x_date(date_breaks = "3 months",  date_labels = "%b %y") +
  theme(legend.position = "right", legend.title = element_text("Shop Types"))
```

In next graph overall trend over time (assuming y ~ x is linear) is graphed used dashed lines.

```{r Shop Location Graph, fig.cap = "Monthly Average Revenue by Shop Location"}
data = train_shop[, .(RevSum = sum(Revenue), ShopAmount = .N, AvgRev = sum(Revenue) / .N), .(LocType, year, month)]
data[, Date := as.Date(paste0(year, "-", month, "-01"))]
ggplot(data, aes(x = Date, y = AvgRev / 1000000, col = LocType)) +
  theme_dt() + labs(x = "Date", y = "Average Revenue (in Million Rubles)", title = "Monthly Average Revenue by Shop Location") +
  geom_smooth(se = FALSE, size = 1, method = "lm", formula = y ~ x, linetype = "dashed") + 
  geom_line(size = 1.4) + 
  scale_x_date(date_breaks = "3 months",  date_labels = "%b %y") +
  theme(legend.position = "right", legend.title = element_text("Location Groups"))
```

```{r Total Revenue Graph, fig.cap = "Monthly Revenue"}
data = train_shop[, .(RevSum = sum(Revenue), ShopAmount = .N, AvgRev = sum(Revenue) / .N), .(year, month)]
data[, Date := as.Date(paste0(year, "-", month, "-01"))]
ggplot(data, aes(x = Date, y = RevSum / 1000000)) +
  theme_dt() + labs(x = "Date", y = "Total Revenue (in Million Rubles)", title = "Monthly Revenue") +
  #geom_smooth(se = FALSE, size = 2) + 
  geom_line(size = 1, col = "orange") + 
  scale_x_date(date_breaks = "3 months",  date_labels = "%b %y")
```

Those graphs hint out many things:

- Within December sales go high up. Most probably due to *Christmas*. A dummy variable may be of need. (Month variable can be used as a factor in R as well.)
- There are many outlying months which has to taken care off very carefully, *i.e. Christmas or Sep 2015 for Web sales*.
- Besides Web sales and *Other* sales, in other words *physical shops in various locations*, have similar trends over time.
- There is an increasing demand for Web-sales and a slightly decreasing demand for non-web related demand over time.
- Outbound Trade & Moscow Sale (Discount?) starts at late 2013 and has a downfall trend. Maybe these sales are replaced with Web sales over time.
- Shops in Moscow have more demand then shops in other locations.
